\vspace{-\baselineskip} Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-2\baselineskip}

### 2017-05-26 papers/paper1/ms.tex
updating simulated data examples using notebooks
rasterizing points on rotation plot
updating tex for new figures etc.

\include{vc}
\include{figures/rotation}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.9\textwidth]{figures/simulated/correct.pdf}
\caption{(left) A simulated dataset.
    (right) The inferred PSD~--~the blue contours encompass 68\% of the
    posterior mass~--~compared to the true PSD (dashed black line).
    \figurelabel{simulated-correct}}
\end{center}
\end{figure}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.9\textwidth]{figures/simulated/wrong-qpo.pdf}
\caption{(left) A simulated dataset.
    (right) The inferred period of the process. The true period is indicated
    by the vertical orange line, the posterior inference using the correct
    model is shown as the blue dashed histogram, and the inference made using
    the ``wrong'' effective model is shown as the black histogram.
    \figurelabel{simulated-wrong}}
\end{center}
\end{figure}

@@

\begin{figure}[p]
\begin{center}
\includegraphics[width=\textwidth]{figures/rotation.pdf}
\caption{Inferred constraints on a quasiperiodic GP model using the covariance
    function in \eq{rot-kernel} and two quarters of \kepler\ data.
(left) The \kepler\ data (black points) and the maximum likelihood model
    prediction (blue curve) for a 50~day subset of the data used.
    The solid blue line shows the predictive mean and the blue contours show
    the predictive standard deviation.
(center) Inferred constraints on the model PSD.
    The dashed line shows the maximum likelihood PSD, the blue solid line
    shows the median of posterior PSD, and the blue contours show the 68\%
    credible region.
(right) Inferred constraints on the model covariance function.
    The dashed line shows the maximum likelihood model, the blue solid line
    shows the median of posterior, and the blue contours show the 68\%
    credible region.
    \figurelabel{rotation}}
\end{center}
\end{figure}

\begin{figure}[p]
\begin{center}
\includegraphics[width=0.5\textwidth]{figures/rotation-period.pdf}
\caption{The posterior constraint on the rotation period of KIC~1430163 using
    the dataset and model from \Figure{rotation}.
    The period is the parameter $P$ in \eq{rot-kernel} and this figure shows
    the posterior distribution marginalized over all other nuisance parameters
    in \eq{rot-kernel}.
    This is consistent with the published rotation period made using the full
    \kepler\ baseline \citep{Mathur:2014}.
    The error bars from the literature are indicated by the two vertical
    dashed lines.
    \figurelabel{rotation-period}}
\end{center}
\end{figure}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/astero-11615890-numax_deltanu_corner.pdf}
\caption{The probabilistic constraints on $\nu_\mathrm{max}$ and $\Delta \nu$
    from the inference shown in \Figure{astero} compared to the published
    value (error bar) based on several years of \kepler\ observations
    \citep{Pinsonneault:2014}.
    \figurelabel{astero-corner}}
\end{center}
\end{figure}

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.9\textwidth]{figures/astero-11615890-comparisons.pdf}
\caption{A comparison between the Lomb-Scargle estimator of the PSD and the
posterior inference of the PSD as a mixture of stochastically-driven simple
harmonic oscillators.
(top) The periodogram of the \kepler\ light curve for KIC~11615890 computed
    on the full four year baseline of the mission. The orange line shows a
    smoothed periodogram and the blue line indicates the level of the
    measurement uncertainties.
(middle) The same periodogram computed using about a month of data.
(bottom) The power spectrum inferred using the mixture of SHOs model described
    in the text and only one month of \kepler\ data.
    The black line shows the median of posterior PSD and the gray contours
    show the 68\% credible region.
    \figurelabel{astero}}
\end{center}
\end{figure}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=\textwidth]{figures/transit-ml.pdf}
    \caption{\emph{(top)} A month-long segment of \kepler\ light curve for
    KIC~1430163 with a synthetic transit model injected (black points) and the
    maximum likelihood model for the stellar variability
    (blue line).
    \emph{(bottom)} The maximum likelihood ``de-trending'' of the data in
    the top panel.
    In this panel, the maximum likelihood model for the stellar variability
    has been subtracted to leave only the transits.
    The de-trended light curve is shown by black error bars and the maximum
    likelihood transit model is shown as a blue line.
    \figurelabel{transit-ml}}
\end{center}
\end{figure}

\begin{figure}[!p]
\begin{center}
\includegraphics[width=\textwidth]{figures/transit-corner.pdf}
\caption{The marginalized posterior constraints on the physical parameters of
    the planet transit in the light curve shown in the top panel of
    \Figure{transit-ml}.
    The two-dimensional contours show the 0.5-, 1-, 1.5, and 2-sigma credible
    regions in the marginalized planes and the histograms along the diagonal
    show the marginalized posterior for each parameter.
    The true values used in the simulation are indicated by blue lines.
    For each parameter, the inference is consistent with the true value.
    \figurelabel{transit-corner}}
\end{center}
\end{figure}

@@

\begin{figure}[tp]
\begin{center}
\includegraphics[width=\textwidth]{figures/benchmark_darwin.pdf}
\caption{A benchmark showing the computational scaling of \celerite\ using the
    \texttt{vanilla} band solver as a function of the number of data points
    and the number of terms.
    \emph{(left)} The cost of computing \eq{gp-likelihood} with a covariance
    matrix given by \eq{celerite-kernel} as a function of the number of data
    points $N$.
    The different lines show the cost for different numbers of terms $J$
    increasing from bottom to top.
    To guide the eye, the straight black line without points
    shows linear scaling in $N$.
    \emph{(right)} The same information plotted as a function of $J$ for
    different values of $N$.
    Each line shows the scaling for a specific value of $N$ increasing from
    bottom to top.
    The black line shows quadratic scaling in $J$.
    \figurelabel{benchmark}}
\end{center}
\end{figure}

\begin{figure}[p]
\begin{center}
\includegraphics[width=\textwidth]{figures/benchmark_darwin_lapack.pdf}
\caption{The same as \Figure{benchmark} but using the \texttt{lapack} solver
    with the \project{MKL} \project{BLAS} library.
    \figurelabel{benchmark-lapack}}
\end{center}
\end{figure}

\begin{figure}[p]
\begin{center}
\includegraphics[width=\textwidth]{figures/benchmark_darwin_sparse.pdf}
\caption{The same as \Figure{benchmark} but using the \texttt{sparse} solver
    from \project{Eigen} \citep{Guennebaud:2010}.
    \figurelabel{benchmark-sparse}}
\end{center}
\end{figure}

>>>>

\include{vc}
\include{figures/rotation/rotation}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.9\textwidth]{figures/simulated/correct.pdf}
\caption{(top left) A simulated dataset (black error bars), and maximum
    likelihood model (blue contours).
    (bottom left) The residuals between the mean predictive model and the data
    shown in the top left figure.
    (right) The inferred PSD~--~the blue contours encompass 68\% of the
    posterior mass~--~compared to the true PSD (dashed black line).
    \figurelabel{simulated-correct}}
\end{center}
\end{figure}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.9\textwidth]{figures/simulated/wrong-qpo.pdf}
\caption{(top left) A simulated dataset (black error bars), and maximum
    likelihood model (blue contours).
    (bottom left) The residuals between the mean predictive model and the data
    shown in the top left figure.
    (right) The inferred period of the process. The true period is indicated
    by the vertical orange line, the posterior inference using the correct
    model is shown as the blue dashed histogram, and the inference made using
    the ``wrong'' effective model is shown as the black histogram.
    \figurelabel{simulated-wrong}}
\end{center}
\end{figure}

@@

\begin{figure}[p]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/rotation/rotation.pdf}
\caption{Inferred constraints on a quasiperiodic GP model using the covariance
    function in \eq{rot-kernel} and two quarters of \kepler\ data.
(top left) The \kepler\ data (black points) and the maximum likelihood model
    prediction (blue curve) for a 60~day subset of the data used.
    The solid blue line shows the predictive mean and the blue contours show
    the predictive standard deviation.
(bottom left) The residuals between the mean predictive model and the data
    shown in the top left figure.
(right) The posterior constraint on the rotation period of KIC~1430163 using
    the dataset and model from \Figure{rotation}.
    The period is the parameter $P$ in \eq{rot-kernel} and this figure shows
    the posterior distribution marginalized over all other nuisance parameters
    in \eq{rot-kernel}.
    This is consistent with the published rotation period made using the full
    \kepler\ baseline \citep{Mathur:2014}.
    The error bars from the literature are indicated by the two vertical
    dashed lines.
    \figurelabel{rotation}}
\end{center}
\end{figure}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/astero/astero.pdf}
\caption{
\response{
(top) The \kepler\ data (black points) and the maximum likelihood model
    prediction (blue curve) for a 10~day subset of the month-long dataset that
    was used for the fit.
    The solid blue line shows the predictive mean and the blue contours show
    the predictive standard deviation.
(bottom) The residuals between the mean predictive model and the data
    shown in the top figure.
}
    \figurelabel{astero}}
\end{center}
\end{figure}

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/astero/astero-corner.pdf}
\caption{The probabilistic constraints on $\nu_\mathrm{max}$ and $\Delta \nu$
    from the inference shown in \Figure{astero} compared to the published
    value (error bar) based on several years of \kepler\ observations
    \citep{Pinsonneault:2014}.
    \figurelabel{astero-corner}}
\end{center}
\end{figure}

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.9\textwidth]{figures/astero/astero-comp.pdf}
\caption{A comparison between the Lomb-Scargle estimator of the PSD and the
posterior inference of the PSD as a mixture of stochastically-driven simple
harmonic oscillators.
(top) The periodogram of the \kepler\ light curve for KIC~11615890 computed
    on the full four year baseline of the mission. The orange line shows a
    smoothed periodogram and the blue line indicates the level of the
    measurement uncertainties.
(middle) The same periodogram computed using about a month of data.
(bottom) The power spectrum inferred using the mixture of SHOs model described
    in the text and only one month of \kepler\ data.
    The black line shows the median of posterior PSD and the gray contours
    show the 68\% credible region.
    \figurelabel{astero}}
\end{center}
\end{figure}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=\textwidth]{figures/transit/transit-ml.pdf}
    \caption{\emph{(top)} A month-long segment of \kepler\ light curve for
    KIC~1430163 with a synthetic transit model injected (black points) and the
    maximum likelihood model for the stellar variability
    (blue line).
    \emph{(bottom)} The maximum likelihood ``de-trending'' of the data in
    the top panel.
    In this panel, the maximum likelihood model for the stellar variability
    has been subtracted to leave only the transits.
    The de-trended light curve is shown by black error bars and the maximum
    likelihood transit model is shown as a blue line.
    \figurelabel{transit-ml}}
\end{center}
\end{figure}

\begin{figure}[!p]
\begin{center}
\includegraphics[width=\textwidth]{figures/transit/transit-corner.pdf}
\caption{The marginalized posterior constraints on the physical parameters of
    the planet transit in the light curve shown in the top panel of
    \Figure{transit-ml}.
    The two-dimensional contours show the 0.5-, 1-, 1.5, and 2-sigma credible
    regions in the marginalized planes and the histograms along the diagonal
    show the marginalized posterior for each parameter.
    The true values used in the simulation are indicated by blue lines.
    For each parameter, the inference is consistent with the true value.
    \figurelabel{transit-corner}}
\end{center}
\end{figure}

@@

\begin{figure}[tp]
\begin{center}
\includegraphics[width=\textwidth]{figures/benchmark_darwin.pdf}
\caption{\dfmtodo{Cholesky} A benchmark showing the computational scaling of \celerite\ using the
    \texttt{vanilla} band solver as a function of the number of data points
    and the number of terms.
    \emph{(left)} The cost of computing \eq{gp-likelihood} with a covariance
    matrix given by \eq{celerite-kernel} as a function of the number of data
    points $N$.
    The different lines show the cost for different numbers of terms $J$
    increasing from bottom to top.
    To guide the eye, the straight black line without points
    shows linear scaling in $N$.
    \emph{(right)} The same information plotted as a function of $J$ for
    different values of $N$.
    Each line shows the scaling for a specific value of $N$ increasing from
    bottom to top.
    The black line shows quadratic scaling in $J$.
    \figurelabel{benchmark}}
\end{center}
\end{figure}
